---
title: "A Machine Learner’s Market: Using Artificial Intelligence to Predict Home Prices"
author: "Lee Waters"
date: "5/9/2021"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# define libraries
library(caret)                                                                  # repeatedcv
library(gbm)                                                                    # varimp
library(reshape2)                                                               # melt correlation matrix
library(ggplot2)                                                                # visualize correlation matrix
library(car)                                                                    # vif
library(Metrics)
```

# Step 0: Load data

```{r}
# load train and test sets
training <- read.csv('train.csv')
testing <- read.csv('test.csv')
```

# Step 1: Pre-processing

Missing values (training)

```{r}
sapply(training, function(x) sum(is.na(x)))                                     # initial missing 

# data imputation
training$LotFrontage[is.na(training$LotFrontage)] <- 0                          # na is rly 0 sq ft 

training$Alley[is.na(training$Alley)] <- 'None'                                 # na is rly none

training <- training[-c(which(is.na(training$MasVnrType))), ]                   # remove 8 missing obs

training$BsmtQual[is.na(training$BsmtQual)] <- 'None'                           # na is rly none
training$BsmtCond[is.na(training$BsmtCond)] <- 'None'                           # na is rly none
training$BsmtExposure[is.na(training$BsmtExposure)] <- 'None'                   # na is rly none
training$BsmtFinType1[is.na(training$BsmtFinType1)] <- 'None'                   # na is rly none
training$BsmtFinType2[is.na(training$BsmtFinType2)] <- 'None'                   # na is rly none

training <- training[-c(which(is.na(training$Electrical))), ]                   # remove 1 missing obs

training$FireplaceQu[is.na(training$FireplaceQu)] <- 'None'                     # na is rly none

training$GarageType[is.na(training$GarageType)] <- 'None'                       # na is rly none
training <- training[, -60]                                                     # remove GarageYrBlt to handle no garage case
training$GarageFinish[is.na(training$GarageFinish)] <- 'None'                   # na is rly none
training$GarageQual[is.na(training$GarageQual)] <- 'None'                       # na is rly none
training$GarageCond[is.na(training$GarageCond)] <- 'None'                       # na is rly none

training$PoolQC[is.na(training$PoolQC)] <- 'None'                               # na is rly none

training$Fence[is.na(training$Fence)] <- 'None'                                 # na is rly none

training$MiscFeature[is.na(training$MiscFeature)] <- 'None'                     # na is rly none

# check for missing values
sapply(training, function(x) sum(is.na(x)))                                     # final missing 
```

Missing values (testing)

```{r}
sapply(testing, function(x) sum(is.na(x)))                                      # initial missing 

# data imputation
temp = table(testing$MSZoning)                                                  # replace w most frequent
names(temp[which.max(temp)])
testing$MSZoning[is.na(testing$MSZoning)] <- 'RL' 
rm(temp)

testing$LotFrontage[is.na(testing$LotFrontage)] <- 0                            # na is rly 0 sq ft

testing$Alley[is.na(testing$Alley)] <- 'None'                                   # na is rly none

temp = table(testing$Utilities)                                                 # replace w most frequent
names(temp[which.max(temp)])
testing$Utilities[is.na(testing$Utilities)] <- 'AllPub' 
rm(temp)

testing$Exterior1st[is.na(testing$Exterior1st)] <- 'other  '                    # assume na is other
testing$Exterior2nd[is.na(testing$Exterior2nd)] <- 'other  '                    # assume na is other

testing$MasVnrType[is.na(testing$MasVnrType)] <- 'None'                         # assume na is none
testing$MasVnrArea[is.na(testing$MasVnrArea)] <- 0                              # assume na is 0

testing$BsmtQual[is.na(testing$BsmtQual)] <- 'None'                             # na is rly none
testing$BsmtCond[is.na(testing$BsmtCond)] <- 'None'                             # na is rly none
testing$BsmtExposure[is.na(testing$BsmtExposure)] <- 'None'                     # na is rly none
testing$BsmtFinType1[is.na(testing$BsmtFinType1)] <- 'None'                     # na is rly none
testing$BsmtFinType2[is.na(testing$BsmtFinType2)] <- 'None'                     # na is rly none
testing$BsmtFinSF1[is.na(testing$BsmtFinSF1)] <- 0                              # assume na is 0
testing$BsmtFinSF2[is.na(testing$BsmtFinSF2)] <- 0                              # assume na is 0
testing$BsmtUnfSF[is.na(testing$BsmtUnfSF)] <- 0                                # assume na is 0
testing$TotalBsmtSF[is.na(testing$TotalBsmtSF)] <- 0                            # assume na is 0
testing$BsmtFullBath[is.na(testing$BsmtFullBath)] <- 0                          # assume na is 0
testing$BsmtHalfBath[is.na(testing$BsmtHalfBath)] <- 0                          # assume na is 0

testing$KitchenQual[is.na(testing$KitchenQual)] <- 'Fa'                         # looked at overall qual and cond

testing$Functional[is.na(testing$Functional)] <- 'Maj2'                         # looked at overall qual and cond

testing$FireplaceQu[is.na(testing$FireplaceQu)] <- 'None'                       # na is rly none

testing$GarageType[is.na(testing$GarageType)] <- 'None'                         # na is rly none
testing$GarageFinish[is.na(testing$GarageFinish)] <- 'None'                     # na is rly none
testing$GarageQual[is.na(testing$GarageQual)] <- 'None'                         # na is rly none
testing$GarageCond[is.na(testing$GarageCond)] <- 'None'                         # na is rly none
testing$GarageCars[is.na(testing$GarageCars)] <- 0                              # assume na is 0
testing$GarageArea[is.na(testing$GarageArea)] <- 0                              # assume na is 0
testing <- testing[, -60]                                                       # remove GarageYrBlt to handle no garage case

testing$PoolQC[is.na(testing$PoolQC)] <- 'None'                                 # na is rly none

testing$Fence[is.na(testing$Fence)] <- 'None'                                   # na is rly none

testing$MiscFeature[is.na(testing$MiscFeature)] <- 'None'                       # na is rly none

testing$SaleType[is.na(testing$SaleType)] <- 'Oth'                              # assume na is other

# check for missing values
sapply(testing, function(x) sum(is.na(x)))                                      # final missing 
```

Prep variables (training)

```{r}
# combine train and test so all factor levels considered
testing$SalePrice <- 0
temp <- rbind(training, testing)

str(temp)                                                                       # initial str

temp$Id <- NULL                                                                 # remove id column

temp$MSSubClass <- as.factor(temp$MSSubClass)

cols_factor <- names(temp[, sapply(temp, class) == 'character'])                # convert chr to factor
temp[cols_factor] <- lapply(temp[cols_factor], factor)
rm(cols_factor)
  
temp$Street <- ordered(temp$Street,
                           levels = c('Grvl', 'Pave'))                          # convert factor to ordinal

temp$LotShape <- ordered(temp$LotShape,
                             levels = c('IR3', 'IR2', 'IR1', 'Reg'))

temp$LandContour <- ordered(temp$LandContour,
                             levels = c('Low', 'HLS', 'Bnk', 'Lvl'))

temp$Utilities <- ordered(temp$Utilities,
                             levels = c('ELO', 'NoSeWa', 'NoSewr', 'AllPub'))

temp$LandSlope <- ordered(temp$LandSlope,
                             levels = c('Sev', 'Mod', "Gtl"))

temp$OverallQual <- ordered(temp$OverallQual)

temp$OverallCond <- ordered(temp$OverallCond)

temp$ExterQual <- ordered(temp$ExterQual,
                             levels = c('Po', 'Fa', 'TA', 'Gd', 'Ex'))

temp$ExterCond <- ordered(temp$ExterCond,
                             levels = c('Po', 'Fa', 'TA', 'Gd', 'Ex'))

##temp$BsmtQual <- ordered(temp$BsmtQual,
##                             levels = c('None', 'Po', 'Fa', 'TA', 'Gd', 'Ex'))

##temp$BsmtCond <- ordered(temp$BsmtCond,
##                             levels = c('None', 'Po', 'Fa', 'TA', 'Gd', 'Ex'))

##temp$BsmtExposure <- ordered(temp$BsmtExposure,
##                             levels = c('None', 'No', 'Mn', 'Av', 'Gd'))

##temp$BsmtFinType1 <- ordered(temp$BsmtFinType1,
##                             levels = c('None', 'Unf', 'LwQ', 'Rec', 'BLQ',
##                                        'ALQ', 'GLQ'))

temp$HeatingQC <- ordered(temp$HeatingQC,
                             levels = c('Po', 'Fa', 'TA', 'Gd', 'Ex'))

temp$CentralAir <- ordered(temp$CentralAir,
                             levels = c('N', 'Y'))

temp$Electrical <- ordered(temp$Electrical,
                             levels = c('Mix', 'FuseP', 'FuseF', 'FuseA',
                                        'SBrkr'))

temp$KitchenQual <- ordered(temp$KitchenQual,
                             levels = c('Po', 'Fa', 'TA', 'Gd', 'Ex'))

temp$Functional <- ordered(temp$Functional,
                             levels = c('Sal', 'Sev', 'Maj2', 'Maj1', 'Mod',
                                        'Min2', 'Min1', 'Typ'))

##temp$FireplaceQu <- ordered(temp$FireplaceQu,
##                             levels = c('None', 'Po', 'Fa', 'TA', 'Gd', 'Ex'))

temp$GarageType <- ordered(temp$GarageType,
                            levels = c('None', 'Detchd', 'CarPort', 'BuiltIn',
                                       'Basment', 'Attchd', '2Types'))

##temp$GarageFinish <- ordered(temp$GarageFinish,
##                             levels = c('None', 'Unf', 'RFn', 'Fin'))

##temp$GarageQual <- ordered(temp$GarageQual,
##                             levels = c('None', 'Po', 'Fa', 'TA', 'Gd', 'Ex'))

##temp$GarageCond <- ordered(temp$GarageCond,
##                             levels = c('None', 'Po', 'Fa', 'TA', 'Gd', 'Ex'))

temp$PavedDrive <- ordered(temp$PavedDrive,
                             levels = c('N', 'P', 'Y'))

str(temp)                                                                       # final str

# separate back train and test
training <- temp[1:1451, ]
testing <- temp[1452:2910, ]
testing$SalePrice <- NULL
rm(temp)
```

# Step 2: Ensemble methods

## RF

```{r, eval=TRUE}
# fit
fitControl <- trainControl(method = "repeatedcv",
                           number = 5,
                           repeats = 5)

start_time <- Sys.time()
set.seed(42)
rfFit1 <- train(SalePrice ~ ., data = training,
                method = "rf",
                trControl = fitControl,
                verbose = FALSE)
end_time <- Sys.time()
end_time - start_time

rfFit1

# variable importance
rfImp <- varImp(rfFit1)
rfImp

plot(rfImp, top = 20)
```

## GBM

```{r, eval=TRUE}
# fit
fitControl <- trainControl(method = "repeatedcv",
                           number = 5,
                           repeats = 5)

start_time <- Sys.time()
set.seed(42)
gbmFit1 <- train(SalePrice ~ ., data = training,
                 method = "gbm",
                 trControl = fitControl,
                 verbose = FALSE)
end_time <- Sys.time()
end_time - start_time

gbmFit1

# variable importance
gbmImp <- varImp(gbmFit1)
gbmImp

plot(gbmImp, top = 20)
```

# Step 3: LM

## EDA and pre-processing

Check for multicollinearity

```{r}
# select features from rf and gbm
names(training[, c(17, 46, 60, 38, 34, 27, 43, 53, 44, 4)])                     # all factors from feature selection
str(training[, c(17, 46, 60, 38, 34, 27, 43, 53, 44, 4)])
str(training[, c(46, 60, 38, 34, 43, 44, 4)])                                   # numerical factors from feature selection

# check correlations                                                            
# see source 1 for reference

# get lower triangle of the correlation matrix
  get_lower_tri<-function(cormat){
    cormat[upper.tri(cormat)] <- NA
    return(cormat)
  }

# get upper triangle of the correlation matrix
  get_upper_tri <- function(cormat){
    cormat[lower.tri(cormat)]<- NA
    return(cormat)
  }

# reorder the correlation matrix
reorder_cormat <- function(cormat){

# Use correlation between variables as distance
dd <- as.dist((1-cormat)/2)
hc <- hclust(dd)
cormat <-cormat[hc$order, hc$order]
}  

# reorder the correlation matrix
cormat = round(cor(training[c(46, 60, 38, 34, 43, 44, 4)]), 2)
cormat <- reorder_cormat(cormat)
upper_tri <- get_upper_tri(cormat)

# melt the correlation matrix
melted_cormat <- melt(upper_tri, na.rm = TRUE)

# create a ggheatmap
ggheatmap <- ggplot(melted_cormat, aes(Var2, Var1, fill = value))+
 geom_tile(color = "white")+
 scale_fill_gradient2(low = "blue", high = "red", mid = "white", 
   midpoint = 0, limit = c(-1,1), space = "Lab", 
    name="Pearson\nCorrelation") +
  theme_minimal()+ # minimal theme
 theme(axis.text.x = element_text(angle = 45, vjust = 1, 
    size = 12, hjust = 1))+
 coord_fixed()

# print the heatmap
print(ggheatmap)
```

Remove `X1stFlrSF` and re-check

```{r}
# reorder the correlation matrix
reorder_cormat <- function(cormat){

# Use correlation between variables as distance
dd <- as.dist((1-cormat)/2)
hc <- hclust(dd)
cormat <-cormat[hc$order, hc$order]
}  

# reorder the correlation matrix
cormat = round(cor(training[c(46, 60, 38, 34, 44, 4)]), 2)
cormat <- reorder_cormat(cormat)
upper_tri <- get_upper_tri(cormat)

# melt the correlation matrix
melted_cormat <- melt(upper_tri, na.rm = TRUE)

# create a ggheatmap
ggheatmap <- ggplot(melted_cormat, aes(Var2, Var1, fill = value))+
 geom_tile(color = "white")+
 scale_fill_gradient2(low = "blue", high = "red", mid = "white", 
   midpoint = 0, limit = c(-1,1), space = "Lab", 
    name="Pearson\nCorrelation") +
  theme_minimal()+ # minimal theme
 theme(axis.text.x = element_text(angle = 45, vjust = 1, 
    size = 12, hjust = 1))+
 coord_fixed()

# print the heatmap
print(ggheatmap)
```

Final features

```{r}
col_all = c(17, 46, 60, 38, 34, 27, 53, 44, 4)
col_continuous = c(46, 38, 34, 44, 4)
col_other = c(17, 27, 53, 60)
```

Check for outliers

```{r}
theme1 <- trellis.par.get()
theme1$plot.symbol$col = rgb(.2, .2, .2, .4)
theme1$plot.symbol$pch = 16
theme1$plot.line$col = rgb(1, 0, 0, .7)
theme1$plot.line$lwd <- 2
trellis.par.set(theme1)
featurePlot(x = training[, col_continuous], 
            y = training$SalePrice, 
            plot = "scatter", 
            layout = c(3, 1))
```

Delete outliers and re-check

```{r}
training <- training[training$GrLivArea < 4500, ]
```

```{r}
theme1 <- trellis.par.get()
theme1$plot.symbol$col = rgb(.2, .2, .2, .4)
theme1$plot.symbol$pch = 16
theme1$plot.line$col = rgb(1, 0, 0, .7)
theme1$plot.line$lwd <- 2
trellis.par.set(theme1)
featurePlot(x = training[, col_continuous], 
            y = training$SalePrice, 
            plot = "scatter", 
            layout = c(3, 1))
```

Visualize categorical predictors

```{r}
par(mfrow = c(2, 2))
boxplot(SalePrice ~ GarageCars, data = training)
boxplot(SalePrice ~ OverallQual, data = training)
boxplot(SalePrice ~ ExterQual, data = training)
boxplot(SalePrice ~ KitchenQual, data = training)
```

Check normality and perform transformations

```{r}
# SalePrice
par(mfrow = c(1, 2))
hist(training$SalePrice)
qqnorm(training$SalePrice, pch = 1, frame = FALSE)
qqline(training$SalePrice, col = "steelblue", lwd = 2)

training$SalePrice <- sqrt(training$SalePrice)

par(mfrow = c(1, 2))
hist(training$SalePrice)
qqnorm(training$SalePrice, pch = 1, frame = FALSE)
qqline(training$SalePrice, col = "steelblue", lwd = 2)
```

```{r}
# GrLivArea
par(mfrow = c(1, 2))
hist(training$GrLivArea)
qqnorm(training$GrLivArea, pch = 1, frame = FALSE)
qqline(training$GrLivArea, col = "steelblue", lwd = 2)

training$GrLivArea <- sqrt(training$GrLivArea)

par(mfrow = c(1, 2))
hist(training$GrLivArea)
qqnorm(training$GrLivArea, pch = 1, frame = FALSE)
qqline(training$GrLivArea, col = "steelblue", lwd = 2)
```

```{r}
# TotalBsmtSF
par(mfrow = c(1, 2))
hist(training$TotalBsmtSF)
qqnorm(training$TotalBsmtSF, pch = 1, frame = FALSE)
qqline(training$TotalBsmtSF, col = "steelblue", lwd = 2)

training$TotalBsmtSF <- sqrt(training$TotalBsmtSF)

par(mfrow = c(1, 2))
hist(training$TotalBsmtSF)
qqnorm(training$TotalBsmtSF, pch = 1, frame = FALSE)
qqline(training$TotalBsmtSF, col = "steelblue", lwd = 2)
```

```{r}
# BsmtFinSF1
par(mfrow = c(1, 2))
hist(training$BsmtFinSF1)
qqnorm(training$BsmtFinSF1, pch = 1, frame = FALSE)
qqline(training$BsmtFinSF1, col = "steelblue", lwd = 2)

training$BsmtFinSF1 <- sqrt(training$BsmtFinSF1)

par(mfrow = c(1, 2))
hist(training$BsmtFinSF1)
qqnorm(training$BsmtFinSF1, pch = 1, frame = FALSE)
qqline(training$BsmtFinSF1, col = "steelblue", lwd = 2)
```

```{r}
# X2ndFlrSF
par(mfrow = c(1, 2))
hist(training$X2ndFlrSF)
qqnorm(training$X2ndFlrSF, pch = 1, frame = FALSE)
qqline(training$X2ndFlrSF, col = "steelblue", lwd = 2)

training$X2ndFlrSF <- sqrt(training$X2ndFlrSF)

par(mfrow = c(1, 2))
hist(training$X2ndFlrSF)
qqnorm(training$X2ndFlrSF, pch = 1, frame = FALSE)
qqline(training$X2ndFlrSF, col = "steelblue", lwd = 2)
```

```{r}
# LotArea
par(mfrow = c(1, 2))
hist(training$LotArea)
qqnorm(training$LotArea, pch = 1, frame = FALSE)
qqline(training$LotArea, col = "steelblue", lwd = 2)

training$LotArea <- sqrt(training$LotArea)

par(mfrow = c(1, 2))
hist(training$LotArea)
qqnorm(training$LotArea, pch = 1, frame = FALSE)
qqline(training$LotArea, col = "steelblue", lwd = 2)
```

## Fit models

Assumptions:

* Linearity (Normal Q-Q plot looks good)
* No autocorrelation (NA)
* Homoscadacisity  (no patterns in Residuals vs. Fitted)
* No outliers (none in Residuals vs. Fitted)
* No influential points (low Cook's Distance)

```{r}
# fit
lmFit1 <- lm(SalePrice ~ OverallQual +
                         GrLivArea +
                         GarageCars +
                         TotalBsmtSF +
                         BsmtFinSF1 +
                         ExterQual +
                         KitchenQual +
                         X2ndFlrSF +
                         LotArea,
                         data = training)

par(mfrow = c(2,2))
plot(lmFit1, which = c(1:4))
```

No multicollinearity of numerical predictors

```{r}
# fit
lmFit2 <- lm(SalePrice ~ GrLivArea +
                         GarageCars +
                         TotalBsmtSF +
                         BsmtFinSF1 +
                         X2ndFlrSF +
                         LotArea,
                         data = training)

vif(lmFit2)
```

Compare with interaction model

```{r}
# fit
lmFit3 <- lm(SalePrice ~ OverallQual *
                         GrLivArea +
                         GarageCars +
                         TotalBsmtSF +
                         BsmtFinSF1 +
                         ExterQual +
                         KitchenQual +
                         X2ndFlrSF +
                         LotArea,
                         data = training)

par(mfrow = c(2,2))
plot(lmFit3, which = c(1:4))
```

Remove influential point and re-fit

```{r}
training <- training[-637, ]
```

```{r}
# fit
lmFit4 <- lm(SalePrice ~ OverallQual *
                         GrLivArea +
                         GarageCars +
                         TotalBsmtSF +
                         BsmtFinSF1 +
                         ExterQual +
                         KitchenQual +
                         X2ndFlrSF +
                         LotArea,
                         data = training)

par(mfrow = c(2,2))
plot(lmFit4, which = c(1:4))
```

Re-fit lmFit1 without influential point

```{r}
# fit
lmFit5 <- lm(SalePrice ~ OverallQual +
                         GrLivArea +
                         GarageCars +
                         TotalBsmtSF +
                         BsmtFinSF1 +
                         ExterQual +
                         KitchenQual +
                         X2ndFlrSF +
                         LotArea,
                         data = training)

par(mfrow = c(2,2))
plot(lmFit5, which = c(1:4))
```

The model with the interaction term (lmFit4) is the better model

```{r}
anova(lmFit5, lmFit4)
```

Train with caret

```{r}
# fit
##fitControl <- trainControl(method = "repeatedcv",
##                           number = 5,
##                           repeats = 5)

start_time <- Sys.time()
set.seed(42)
lmFit6 <- train(SalePrice ~ OverallQual *
                            GrLivArea +
                            GarageCars +
                            TotalBsmtSF +
                            BsmtFinSF1 +
                            ExterQual +
                            KitchenQual +
                            X2ndFlrSF +
                            LotArea,
                            data = training,
                            method = "lm",
                            trControl = fitControl,
                            verbose = FALSE)
end_time <- Sys.time()
end_time - start_time

lmFit6
```

# Step 4: Testing

|     | RMSE     | Rsquared  | MAE      |
|-----|----------|-----------|----------|
| RF  | 29256.82 | 0.8677282 | 17351.59 |
| GBM | 28553.05 | 0.8720533 | 16923.64 |
| LM  | 71.75269 | 0.7058348 | 24.21299 |

```{r}
# predict from best model
gbmPreds1 <- predict(gbmFit1, newdata = testing)
```

```{r}
# write to csv
temp <- read.csv('test.csv')
id <- temp$Id
rm(temp)

lmw_submission <- data.frame(Id = id, SalePrice = gbmPreds1)
write.csv(lmw_submission, file = 'lmw_submission.csv', row.names = FALSE)
```

# References
ggplot2: Quick correlation matrix heatmap - R software and data visualization - Easy Guides - Wiki - STHDA. Sthda.com. Published 2020. Accessed March 30, 2021. http://www.sthda.com/english/wiki/ggplot2-quick-correlation-matrix-heatmap-r-software-and-data-visualization